---
title: "echartr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{echartr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r echartr, include=FALSE}
library(echartr)
```

The goal of `echartr` is to provide an R interface to [Echarts](https://echarts.apache.org) that is as close to the original JavaScript API as possible. 

## Creating a chart

The function for creating a new chart is `echartr()`, which takes an `option` argument. This argument is a list of lists that defines the chart's configuration.

For example, we can replicate [this basic line chart](https://echarts.apache.org/examples/en/editor.html?c=line-simple) like so:

```{r}
echartr(option = list(
  xAxis = list(type = "category", data = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
  yAxis = list(type = "value"),
  series = list(type = "line", data = c(150, 230, 224, 218, 135, 147, 260))
))
```
## Working with data frames

Since most data being visualised in R come from a data frame, `echartr` provides helper functions for building the `option` object from data frames. These functions are prefixed with `ec_`. 

Building the same chart using `ec_line()` looks like:

```{r}
data <- tibble::tibble(
  day = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
  value = c(150, 230, 224, 218, 135, 147, 260)
)

echartr(option = list(
  xAxis = list(type = "category"),
  yAxis = list(type = "value"),
  series = ec_line(data, day, value)
))
```

All `ec_` functions take a data frame as the first argument. `ec_line` expects a `y` argument and optionally an `x` argument, which can be **any expression** that can be evaluated from the data frame columns and local environment. 

For example, we can transform the y-values within the `ec_line` arguments:

```{r}
inverse <- function(x) {
  1 / x
}

echartr(option = list(
  xAxis = list(type = "category"),
  yAxis = list(type = "value"),
  series = ec_line(data, day, inverse(value))
))
```
## Series properties

Other [series properties](https://echarts.apache.org/en/option.html#series) can be provided as named arguments to the `ec_` functions. For example, specifying a `lineStyle`:

```{r}
echartr(option = list(
  xAxis = list(type = "category"),
  yAxis = list(type = "value"),
  series = ec_line(data, day, value, lineStyle = list(width = 5, type = "solid"))
))
```
These can also refer to the data frame columns - any series property taking multiple values will result in multiple series. Consider the case where we want to plot 2 different weeks:

```{r}
data2 <- data |>
  dplyr::mutate(week = "Week 1") |>
  dplyr::bind_rows(
    tibble::tibble(
      day = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
      value = c(180, 200, 252, 177, 200, 212, 276),
      week = "Week 2"
    )
  )

echartr(option = list(
  xAxis = list(type = "category"),
  yAxis = list(type = "value"),
  legend = list(show = TRUE),
  series = ec_line(data2, day, value, name = week)
))
```
This works even for deeply nested properties:

```{r}
echartr(option = list(
  xAxis = list(type = "category"),
  yAxis = list(type = "value"),
  legend = list(show = TRUE),
  series = ec_line(
    data2, day, value, 
    name = week, 
    label = list(
      show = (week == "Week 1") # Show label for Week 1 only
    )
  )
))
```
## Debugging in Javascript

`echartr` is designed based on the belief that using the Echarts Javascript API is the most straightforward way to build and understand Echarts. 

That's why it's always easy to see the javascript equivalent of your R code - printing the output of any of the `ec_` functions will show you the equivalent javascript.

```{r}
ec_line(data2, y = value, name = week)
```
For an `option` list-tree, using `new_option` as the outermost `list` will do the same:

```{r}
new_option(
  xAxis = list(type = "category", data = unique(data$day)),
  yAxis = list(type = "value"),
  series = ec_line(data, y = value)
)
```
These can be written straight to the clipboard with `ec_clip` for easy pasting into the [Echarts online editor](https://echarts.apache.org/examples/en/editor.html):

```{r, eval = FALSE}
ec_clip(ec_line(data2, y = value, name = week))
```

## Next steps

This guide has shown you how to create a basic chart using `echartr` - to get the most out of `echartr`, check out the guides on [events and actions](#) and [shiny bindings](#)
