---
title: "Shiny"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(echartr)
library(shiny)
library(rlang) # for `%||%`
```

# Rendering

The ui function is `echartrOutput()`.

The server function is `renderEchartr()`.

```{r, eval = FALSE}
library(shiny)

shinyApp(
  ui = fluidPage(
    titlePanel("Echarts in Shiny"),
    echartrOutput("chart")
  ),
  server = function(input, output) {
    output$chart <- renderEchartr({
      echartr(option = list(
        xAxis = list(type = "value"),
        yAxis = list(type = "value"),
        series = ec_scatter(iris, 
          x = Petal.Width, 
          y = Petal.Length, 
          name = as.character(Species)
        )
      ))
    })
  }
)
```

# Updating

By default, on each render the chart will be fully disposed of and recreated.

To override this behavior and update the existing chart, use `update = TRUE` in `echartr()`:

```{r, eval = FALSE}

# Toggle between two series to show the difference between updating and rerendering
series <- list(
  ec_scatter(iris, x = Petal.Width,y = Petal.Length, name = Species),
  ec_scatter(iris,x = Sepal.Width, y = Sepal.Length, name = Species)
)

shinyApp(
  ui = fluidPage(
    titlePanel("Echarts in Shiny"),
    echartrOutput("chart_replace"),
    echartrOutput("chart_update")
  ),
  server = function(input, output) {
    
    # Toggle the series every 2 seconds
    rv <- reactiveValues(series_index = 2L)
    current_series <- reactive({
      invalidateLater(2000L)
      
      isolate({
        rv$series_index <- 3L - rv$series_index
        series[[rv$series_index]]
      })
    })
    
    # Chart with default update = FALSE
    output$chart_replace <- renderEchartr({
      echartr(option = list(
        title = list(text = "Replace"),
        xAxis = list(type = "value"),
        yAxis = list(type = "value"),
        series = current_series()
      ))
    })
    
    # Chart with update = TRUE
    output$chart_update <- renderEchartr({
      echartr(update = TRUE, option = list(
        title = list(text = "Update"),
        xAxis = list(type = "value"),
        yAxis = list(type = "value"),
        series = current_series()
      ))
    })
  }
)
```


The other way to update an existing chart is with `updateEchartr()` within an `observe()`.

Updating the `option` argument this way will avoid the `recalculating` state.

Another use of `updateEchartr()` is to dispatch actions. For example, updating which series is highlighted based on a selectInput:

```{r, eval = FALSE}
series <- ec_scatter(
  iris, Petal.Length, Petal.Width, 
  name = as.character(Species),
  selectedMode = "multiple",
  silent = TRUE,
  itemStyle = list(opacity = 0.3),
  emphasis = list(
    itemStyle = list(opacity = 1)
  )
)

shinyApp(
  ui = fluidPage(
    titlePanel("Echarts in Shiny"),
    sidebarLayout(
      sidebarPanel(
        selectInput(
          "highlight", "Highlight: ", 
          choices = as.character(iris$Species),
          selected = 1
        )
      ),
      mainPanel(
        echartrOutput("chart")
      )
    )
  ),
  server = function(input, output) {
    
    output$chart <- renderEchartr({
      echartr(
        option = list(
          xAxis = list(type = "value"),
          yAxis = list(type = "value"),
          series = series
        ),
        dispatch = list(
          # Highlight initial selection immediately
          list(type = "highlight", seriesName = isolate(input$highlight))
        )
      )
    })
    
    # Update highlight based on input
    observe({
      updateEchartr(outputId = "chart", dispatch = list(
        list(
          type = "highlight", 
          seriesName = input$highlight
        ),
        list(
          type = "downplay", 
          seriesName = setdiff(as.character(iris$Species), input$highlight)
        )
      ))
    })
  }
)
```

# Event listeners

Either `renderEchartr()` or `updateEchartr()` can be used to specify which event data to send to the server, via the `listen` parameter.

To de-register event listeners, you can use the `off` parameter of `updateEchartr()`

```{r, eval = FALSE}
listeners <- c("click", "mouseover", "mouseout", "selectchanged", "legendselectchanged")

shinyApp(
  ui = fluidPage(
    titlePanel("Echarts in Shiny"),
    sidebarLayout(
      sidebarPanel(
        selectInput(
          "listen_to", "Listen to: ", 
          choices = listeners,
          multiple = TRUE
        ),
        h6("Latest event:"),
        verbatimTextOutput("events")
      ),
      mainPanel(
        echartrOutput("chart")
      )
    )
  ),
  server = function(input, output) {
    
    output$chart <- renderEchartr({
      echartr(option = list(
        xAxis = list(type = "value"),
        yAxis = list(type = "value"),
        legend = list(show = TRUE),
        series = ec_scatter(
          iris, Petal.Length, Petal.Width, 
          name = as.character(Species),
          selectedMode = "multiple"
        )
      ))
    })
    
    rv <- reactiveValues(
      listen_to = list(), # Track how listeners change
      latest_event = NULL
    )

    # Register which events to listen for
    observe({
      listen_to <- input$listen_to %||% character()
      
      updateEchartr(
        outputId = "chart",
        listen = setdiff(listen_to, rv$listen_to), # Add new listeners
        off = setdiff(rv$listen_to, listen_to) # Remove old listeners
      )
      
      rv$listen_to <- listen_to
    }) |>
      bindEvent(input$listen_to, ignoreNULL = FALSE)
    
    # Track the latest event
    observeEvent(input$chart_click, {
      rv$latest_event <- input$chart_click
    })
    observeEvent(input$chart_mouseover, {
      rv$latest_event <- input$chart_mouseover
    })
    observeEvent(input$chart_mouseout, {
      rv$latest_event <- input$chart_mouseout
    })
    observeEvent(input$chart_selectchanged, {
      rv$latest_event <- input$chart_selectchanged
    })
    observeEvent(input$chart_legendselectchanged, {
      rv$latest_event <- input$chart_legendselectchanged
    })
    
    # Display the latest event
    output$events <- renderPrint({
      if (is.null(rv$latest_event)) "No events yet" else rv$latest_event
    })
  }
)
```
